---
title: "SimAnneal"
author: "Paul M"
date: "1/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here's some pseudocode for the simulated annealing [SA] algorithm. Your job is to fill in the gaps.


First, here's a function to practice minimizing. The minimum is at (0,0)
```{r}
XYSquared<-function(x)
{
  return ((x[1]^2+x[2]^2))
}
```

Simulated annealing is a random algorithm, so unless you control the random 
number generator you will get different answers each time you run it. That 
makes life hard when you are debugging, so we take care of that by setting 
the 'seed' for the random number generator. This means we will get the same 
sequencxe of random numbers each time we run the code.

```{r}
set.seed(345)   # don't forget to set the random nuber generator seed
```

Now, here's the outline of a funcrtion that implements SA.
Once complete, you will pass this function these arguments:
* The name of the function you are trying to minimize ("Fn")
* The starting coordinates ("StartPoint")
* The initial temperature ("InitialTemp")
* The final temperature ("FinalTemp")
* The proportion by which to decrease the temperature each iteration ("TempDecreaseRate")

```{r SA}
SimAnneal2D<-function(Fn,StartPoint,InitialTemp,FinalTemp,TempDecreaseRate){
  StepSize<-0.5   # how much are we going to move by for each step?
  X<-StartPoint[1]
  Y<-StartPoint[2]
  # record the value the function takes at this point (call it FnVal, say)
  # ADD code to calculate FnVal here

  # Set the temperature (Temp) to = InitialTemp,
  # ADD CODE to set TEMP here

  plot(X,Y,xlim=c(-2*X,2*X),ylim=c(-2*Y,2*Y)) # set up a plot
  points(0,0,pch=19,col="red")  # plot the Minimum - this is where you are hoping to end up
  while (Temp>FinalTemp){  # we keep going until things cool down
    #propose new point
    NewX<-X+rnorm(1,mean=0,sd=StepSize)  # We use a random walk in which we add a normal rv with mean 0 to each coordinate
    NewY<-Y+rnorm(1,mean=0,sd=StepSize)
    New<-c(NewX,NewY)
    #decide whether to move
    NewVal<- Fn(New)  # the value the function takes at the new point

    # Calculate the value of H
    # ADD CODE FOR H HERE

    p<-runif(1)  # the random number that is going to help us decide whether to move
    if (p<h){ # we move
       # add an arrow to the plot showing where we moved:
      arrows(X,Y,NewX,NewY,length=0.05)
      # update your records of where we are:...X<-NewX   Y<-NewY
      # Add code here

      # pause for a bit - otherwise the plots flash by too quickly to see properly
      Start.Time<-Sys.time()
      while (Sys.time()<Start.Time+0.02){}   # there is probably a better way of doing this
      # plot the path you are taking
      points(X,Y,pch='.',col="blue")

      # update the record of the function value:
      FnVal<-NewVal
    }
    #reduce temperature
    Temp<-Temp*(1-TempDecreaseRate)

  }
  points(X,Y,pch=19,col="blue")   # a big blue point to show where we ended up
  return (c(X,Y,Fn(X,Y)))  # return the coordinates of our final resting place, and the function value
}
```

Call the function like this (for this example we start at the point (5,5), with a temperature going from 10 to 0.1)
```{r test}
SimAnneal2D(XYSquared,c(5,5),10,0.1,0.02)
```
